Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-11-27 leadership_across_cultures_contexts.Rmd

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most frequently noted in the ethnographic record.

<!-- At the text record level, hierarchical cluster analysis was then used to identify natural groupings of all our coded variables. If variables from a model clustered together, and separately from the other variables, that would lend support to the theoretical cohesiveness of that model. If, on the other hand, variables from different models clustered together, that would suggest theoretical models may need revision or that clusters had identified components of a novel theoretical model.  -->

To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` function variables, all of which take binary values (0, 1), we used logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable. We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors. Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitative measure of leader qualities and leader functions for additional exploratory analyses.

To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015]. These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.

@@

>>>>

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.

Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together. Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001]. Or they might suggest new and theoretically important dimensions of leadership.

There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].

Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables. We optimized rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), using cross-validation, per recommendation of the package authors.

<!-- To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015]. These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary. -->

@@

We also performed hierarchical cluster analysis on our quality and function variables. We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.

Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size. We used linear mixed effects models from the lme4 package [@xx], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts. 
>>>>>>> 59ccacb71b5cc33ebe98b5790a61cccb17f65052
